---
title: 'Snippets: data frame manipulation'
author: ''
date: '2018-02-25'
slug: snippets-data-frame-manipulation
categories:
  - R
tags:
  - dataframe
summary: "These commands are relevant for data frames and their manipulation."
output:
  blogdown::html_page:
    toc: true
---

```{r , setup, include = FALSE}
library(here)
library(tidyverse)
library(rvest)

url <- "https://www.fangraphs.com/leaders.aspx?pos=all&stats=bat&lg=all&qual=y&type=8&season=2018&month=0&season1=2018&ind=0"

url %>%
  read_html() %>%
  html_nodes(xpath = '//*[@id="LeaderBoard1_dg1_ctl00"]') %>%
  html_table() %>%
  .[[1]] %>% # pull out the first member of the list
  as.data.frame() %>%
  setNames(.[2,]) %>% # identify colnames by row
  slice(-c(1:3)) %>%
  write.csv(here("data", "mlb.csv"), row.names = FALSE)

mlb <- read.csv(here("data", "mlb.csv"))

```

Snippet files are periodically updated with how-to's for data wrangling. These commands are relevant for data frames and their manipulation. 

#  Multiple character columns to numeric
  
```{r eval = FALSE}
# can either reference columns by position or names
df[c(1,3,5,9)] <- sapply(df[c(1,3,5,9)], as.numeric)
```

#  Data frame factor to character 

```{r eval = FALSE}
df <- df %>% mutate_if(is.factor, as.character)

# this can also be done using the purrr package 
library(purrr)
library(dplyr)
df <- df %>% map_if(is.factor, as.character) %>% as_data_frame
```

#  Renaming columns with special characters

This is primarily useful if special characters are necessary for graphing. For instance, if we were to use the `read.csv` command when reading in a csv that has special characters such as a `%` or `+` sign, R converts these to dots `(.)`. But if we were to try to use these variables in graphs, we need to put these characters back into the variable names.  
```{r eval = TRUE}
# can also reference with specific columns using names(df)[i:j]
names(mlb)
# choose which cols to name change
names(mlb)[1:17] <- gsub(x = names(mlb)[1:17], pattern = "\\.", replacement = "%")
# now can add in the + to the final column
names(mlb) <- gsub(x = names(mlb), pattern = "\\.", replacement = "+")
names(mlb)
```

# Setting NAs to Zero 

Assume a data frame that has NAs and you want to set any NA to zero, it is straightforward to do so. 
```{r eval = TRUE}

df = tibble(
  x = c(1,NA,3,NA,5), 
  y = c(1:5), 
  z = x ^ 2 + y) %>% 
  data.frame()

df[is.na(df)] <- 0

head(df)
```

We can also selectively choose the column to change by referencing it specifically 
```{r eval = TRUE}
df = tibble(
  x = c(1,NA,3,NA,5), 
  y = c(1,NA,3,NA,5),
  z = 1:5) %>%
  data.frame()

df$x[is.na(df$x)] <- 0

head(df)
```

Or, within a dplyr pipe, we can use this, and this will work across the entire data frame 
```{r eval = FALSE}
mutate_all(funs(replace(., is.na(.), 0)))
```

# Replacing NAs with Previous Value by Group

Occasionally you can end up with a data frame that has a series of NA, such as below. The goal is to replace those NA simply. The ```tidyr``` package can handle this with the ```fill``` command. Note that this is a slow process though, so if you have a large dataframe, this is probably not the best option. 
``` {r eval = TRUE}
df <- data.frame(userID = c(rep(21, 10), rep(22, 10)), 
                 date = c("2017-12-10", rep(NA, 9), 
                          rep("2017-12-11", 10)))
df

df %>% group_by(userID) %>% 
  fill(date)
```


