---
title: 'Snippets: data frame manipulation'
author: ''
date: '2018-02-25'
slug: snippets-data-frame-manipulation
categories:
  - R
tags:
  - dataframe
---

```{r , setup, include = FALSE}
library(here)
library(tidyverse)
mlb <- read_csv(here('data', 'mlb.csv'))
```

Snippet files are periodically updated with how-to's for data wrangling. These commands are relevant for data frames and their manipulation. 

### Multiple character columns to numeric
```{r eval = FALSE}
# can either reference columns by position or names
df[c(1,3,5,9)] <- sapply(df[c(1,3,5,9)], as.numeric)
```

### Data frame factor to character 
```{r eval = FALSE}
df <- df %>% mutate_if(is.factor, as.character)

# this can also be done using the purrr package 
library(purrr)
library(dplyr)
df <- df %>% map_if(is.factor, as.character) %>% as_data_frame
```

### Renaming columns with special characters
This is primarily useful if special characters are necessary for graphing. For instance, when reading in a csv that has special characters such as a `%` or `+` sign, R converts these to dots `(.)`. But if we were to try to use these variables in graphs, we need to put these characters back into the variable names.  
```{r eval = TRUE}
# can also reference with specific columns using names(df)[i:j]
names(mlb)
# choose which cols to name change
names(mlb)[1:17] <- gsub(x = names(mlb)[1:17], pattern = "\\.", replacement = "%")
# now can add in the + to the final column
names(mlb) <- gsub(x = names(mlb), pattern = "\\.", replacement = "+")
names(mlb)
```

### Setting NAs to Zero 
Assume a data frame that has NAs and you want to set any NA to zero, it is straightforward to do so. 
```{r eval = TRUE}
df <- tibble(x = c(1,NA,3,NA,5), 
                 y = c(1:5),
                 z = x ^ 2 + y) %>%
  data.frame()
df[is.na(df)] <- 0
head(df)
```

We can also selectively choose the column to change by referencing it specifically 
```{r eval = TRUE}
df <- tibble(x = c(1,NA,3,NA,5), 
                 y = c(1,NA,3,NA,5),
                 z = 1:5) %>%
  data.frame()
df$x[is.na(df$x)] <- 0
head(df)
```

Or, within a dplyr pipe, we can use this, and this will work across the entire data frame 
```{r eval = FALSE}
mutate_all(funs(replace(., is.na(.), 0)))
```

### Replacing NAs with Previous Value by Group
Occasionally you can end up with a data frame that has a series of NA, such as below. The goal is to replace those NA simply. The ```tidyr``` package can handle this with the ```fill``` command. Note that this is a slow process though, so if you have a large dataframe, this is probably not the best option. 
``` {r eval = TRUE}
df <- data.frame(userID = c(rep(21, 10), rep(22, 10)), 
                 date = c("2017-12-10", rep(NA, 9), 
                          rep("2017-12-11", 10)))
df

df %>% group_by(userID) %>% 
  fill(date)
```


