---
title: 'Snippets: General R'
author: ''
date: '2018-04-23'
slug: snippets-r-general
categories: 
- R
tags:
- R
- RStudio
summary: "These functions and commands are relevant for working with R and RStudio."
---

Snippet files periodically updated. These functions and commands are relevant for working with R and RStudio. 

# Easily updating R

The `updateR` package is fantastic. Just using this line of code with the admin_password set will update everything and keep all packages in the new version of R. 

```{r, eval = FALSE}
updateR(admin_password = "admin_password")
```

# Finding all functions in a package

This is especially useful when using a dev version of a package. Often, there are github files, but it's sometimes easier to just use `asNamespace` to see functions. 

For example, let's look at the functions available within the spotifyR package. 

```{r, echo = TRUE}
library(spotifyr)
fns <- unclass(lsf.str(envir = asNamespace("spotifyr"), all = T))

fns
```

# Loading all functions within a folder

The `R.utils` package also has a command for this, but this is a simple way to source all of your functions at once. 

```{r eval = FALSE}
# set directory location, or use here() rather than glue
# load functions
file.sources = list.files(path = glue(dir, "functions"), pattern="*.R")
sapply(glue("{dir}functions/{file.sources}"), source, .GlobalEnv)
```




